Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-04-10 docs/ACM/paper.tex
Fixing caption of Figure 13.
Adding Description of optimizations.

\noindent
\textbf{Benchmarking:}
we have used different benchmarks in our experiments.
In Sections~\ref{sub:performance} and~\ref{sub:optimizations} we use
\textsf{BSC-Bots}\footnote{\url{https://github.com/bsc-pm/bots/tree/master/serial}}~\cite{Duran09} and
\textsf{Swan}\footnote{\url{http://cuda.dcc.ufmg.br/swan/}}~\cite{Moreira17}.
In Section~\ref{sub:versatility} we use the LLVM test suite, and in
Section~\ref{sub:scalability} we use random programs produced with
\textsf{CSmith}~\cite{Yang11}.
Runtime numbers are averages of five executions.
We have not changed manually any of the benchmarks.

%PERFORMANCE
%Here, we show the results for our small benchmarks (toys & bots)
\subsection{Performance}
\label{sub:performance}

Figure~\ref{fig:TM_Performance} compares the runtime of programs produced by
\Taskminer{} against either the original -- sequential -- program (the baseline,
compiled with \textsf{clang} -O3)
or versions of said programs annotated manually.
In this experiment we use the two benchmarks mentioned in the beginning of this
Section: \textsf{BSC-Bots}~\cite{Duran09} (\textsf{fft} to \textsf{jacobi}
in Fig.~\ref{fig:TM_Performance}) and \textsf{Swan}~\cite{Moreira17}
(\textsf{dfs} to \textsf{private}).
Both these benchmarks come with sequential and parallel versions.

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\columnwidth]{images/TM_Performance}
\caption{Speedup comparisons -- automatic annotations compare favourably against
manual interventions, and lead to great performance gains.
The Y-axis shows the speedup of either manual intervention or the \Taskminer{}
(this paper) upon the original programs, in number of times.
Small numbers in boxes show speedup achieved by \Taskminer.}
\label{fig:TM_Performance}
\end{center}
\end{figure}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=1\columnwidth]{images/Stub}
\caption{The benefit of our optimizations.
Each optimization (cost model and recursion bounds) lead to
non-trivial speedups.}
\label{fig:Opt_Speedups}
\end{center}
\end{figure}

In most of the benchmarks, the presence of task cost estimates made little difference in performance of
the automatically annotated programs. In some of them, the cost model was very decisive to 
guarantee a speed up against the sequential versions, though. These can be seen in Figure \ref{fig:Opt_Speedups}.
A coarse gain in speed up was noticed in the \textit{nqueens.c} program, in which the recursion bounding
avoided the creation of smaller and exponentially increasing tasks, which caused a significant speed
up in its execution. 
If on one hand, the cost model is able to avoid the creation and the dispatch of tasks that are too small
and are not going to pay off for the runtime cost, on the other hand computing the cost for tasks can add
some overhead to the programs. This is easily seen in the \textit{bellmanford} program. For larger
matrices, larger is the number of paths to be computed, so in this case, {\Taskminer} will try to
compute the cost for every task, which we believe caused this minor slowdown. However, it's
clear that the cost model greatly improved {\Taskminer}.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=1\columnwidth]{images/TM_GeneralProgs}
\caption{20 best speedups obtained by \Taskminer{} when applied onto the LLVM
test suite.
The larger the bar, the better.
\textbf{LoC} stands for ``Lines of C".}
\label{fig:TM_GeneralProgs}
\end{center}
\end{figure}

>>>>

\noindent
\textbf{Benchmarking:}
we have used different benchmarks in our experiments.
In Sections~\ref{sub:performance} and~\ref{sub:optimizations} we use
\textsf{BSC-Bots} (\url{https://github.com/bsc-pm/bots/tree/master/serial})~\cite{Duran09} and
\textsf{Swan} (\url{http://cuda.dcc.ufmg.br/swan/})~\cite{Moreira17}.
In Section~\ref{sub:versatility} we use the LLVM test suite, and in
Section~\ref{sub:scalability} we use random programs produced with
\textsf{CSmith}~\cite{Yang11}.
Runtimes are averages of five executions.
No benchmark has been changed manually.

\subsection{Performance}
\label{sub:performance}

\begin{figure}[b!]
\begin{center}
\includegraphics[width=1\columnwidth]{images/TM_Performance}
\caption{Speedup comparisons -- automatic annotations compare favourably against
manual interventions, and lead to great performance gains.
The Y-axis shows the speedup of either manual intervention or the \Taskminer{}
(this paper) upon the original programs, in number of times.
Small numbers in boxes show speedup achieved by \Taskminer.}
\label{fig:TM_Performance}
\end{center}
\end{figure}

Figure~\ref{fig:TM_Performance} compares the runtime of programs produced by
\Taskminer{} against either the original program (the baseline,
compiled with \textsf{clang} -O3)
or versions of said programs annotated manually.
In this experiment we use the benchmarks \textsf{BSC-Bots}~\cite{Duran09} (\textsf{fft} to \textsf{jacobi}
in Fig.~\ref{fig:TM_Performance}) and \textsf{Swan}~\cite{Moreira17}
(\textsf{dfs} to \textsf{private}).
Both these benchmarks come with sequential and parallel (manually
annotated) versions.

@@

\begin{figure}[b!]
\begin{center}
\includegraphics[width=1\columnwidth]{images/Optimizations}
\caption{Benefit of task pruning (Secs.~\ref{sub:profit} and~\ref{sub:ir}).
\textbf{RFC}: tasks created within \underline{R}ecursive \underline{F}unction
\underline{C}alls.
\textbf{NRC}: interprocedural tasks created around
\underline{N}on-\underline{R}ecursive function \underline{C}alls.
\underline{Reg}: tasks involving \underline{Reg}ions without function calls.
}
\label{fig:Optimizations}
\end{center}
\end{figure}

This paper describes two ways to optimize task placement.
Both these techniques are based on the idea of ``task pruning": we avoid
creating tasks if we deem them unprofitable.
The first technique prunes tasks judged unprofitable by the cost-model of
Section~\ref{sub:profit}; the second prunes tasks that are too deep in the
recursion stack, as described in Section~\ref{sub:ir}.
Figure~\ref{fig:Optimizations} illustrates the benefits of these two techniques.
We show the gains -- in terms of speedup over the baseline -- of the three
benchmarks that
benefit the most from each type of task pruning, among those earlier seen in
Figure~\ref{fig:TM_Performance}.

In three benchmarks, \textsf{private}, \textsf{boruvka} and
\textsf{bellman-ford}, the cost model avoids creating tasks around loops that
initialize data-structures.
For instance, the following loop, in line 75 of \textsf{bellmanford.c}, would be
parallelized by \Taskminer, were it not for the cost-model marking it as
unprofitable:

\begin{verbatim}
  for (long unsigned i = 0; i < N; i++)
    for (long unsigned j = 0; j < N; j++)
      *(G + i * N + j) = rand();
\end{verbatim}

Similarly, recursion bounding (Sec.~\ref{sub:ir}), although
simple, is effective in eliminating the excessive number of tasks.
Figure~\ref{fig:Optimizations} shows the effect of this optimization upon
three \textsf{BSC-Bots} benchmarks: \textsf{health}, \textsf{nqueens} and
\textsf{strassen}.
These programs were designed to illustrate the parallelization of
divide-and-conquer algorithms~\cite{Duran09}, and contain a large number of
recursive calls (see \textbf{RFC}) in Figure~\ref{fig:Optimizations}.
Pruning at higher levels of the recursion tree lets them deliver
non-negligible speedups onto the baseline programs.
Were pruning absent, then we would observe slowdowns in \textsf{health}
and \textsf{nqueens}.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=1\columnwidth]{images/TM_GeneralProgs}
\caption{Speedups obtained by \Taskminer{} when applied onto the LLVM
test suite.
The larger the bar, the better.
\textbf{LoC} stands for ``Lines of C".}
\label{fig:TM_GeneralProgs}
\end{center}
\end{figure}
